---
title: "Assignment 3"
editor: visual
---

```{r}
#| echo: false
#| message: false
#| warning: false

library(tidycensus)
library(tigris)
library(sf)
library(dplyr)
library(ggplot2)
library(readr)

options(tigris_use_cache = TRUE)

state_abbr <- "TX"
geo_level <- "county"

my_vars <- c(
  income = "B19013_001",
  pov = "B17001_002",
  povden = "B17001_001"
)

year_acs <- 2024
survey <- "acs5"

acs_wide <- get_acs(
  geography = geo_level,
  variables = my_vars,
  state = state_abbr,
  year = year_acs,
  survey = survey,
  geometry = TRUE,
  output = "wide"
)

acs_wide <- acs_wide |>
  mutate(
    pov_rate = povE / povdenE,
    pov_moe = moe_prop(povE, povdenE, povM, povdenM)
  )
```

## Median Household Income Map

```{r}
ggplot(acs_wide) +
  geom_sf(aes(fill = incomeE), color = NA) +
  scale_fill_viridis_c(
    name = "Median HH Income"
  ) +
  labs(
    title = "Median Household Income in Texas",
    subtitle = "2024 ACS 5-Year Estimates by County",
    caption = "Source: U.S. Census Bureau via tidycensus"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.caption = element_text(size = 8),
    legend.position = "right"
  )
```

## Poverty Rate: Top and Bottom 10 Counties

```{r}
top10 <- acs_wide |>
  st_drop_geometry() |>
  arrange(desc(pov_rate)) |>
  select(NAME, pov_rate, pov_moe) |>
  slice_head(n = 10)

bottom10 <- acs_wide |>
  st_drop_geometry() |>
  arrange(pov_rate) |>
  select(NAME, pov_rate, pov_moe) |>
  slice_head(n = 10)

top10_table <- top10 |>
  mutate(
    pov_rate = sprintf("%.2f%%", pov_rate * 100),
    pov_moe = sprintf("±%.2f pp", pov_moe * 100)
  ) |>
  rename(
    County = NAME,
    `Poverty Rate` = pov_rate,
    `MOE` = pov_moe
  )

bottom10_table <- bottom10 |>
  mutate(
    pov_rate = sprintf("%.2f%%", pov_rate * 100),
    pov_moe = sprintf("±%.2f pp", pov_moe * 100)
  ) |>
  rename(
    County = NAME,
    `Poverty Rate` = pov_rate,
    `MOE` = pov_moe
  )

knitr::kable(
  top10_table,
  caption = "Texas Counties with the Highest Poverty Rates"
)

knitr::kable(
  bottom10_table,
  caption = "Texas Counties with the Lowest Poverty Rates"
)
```

## Uncertainty Check

```{r}
uncertainty_pair <- acs_wide |>
  st_drop_geometry() |>
  filter(NAME %in% c("Dimmit County, Texas", "Zapata County, Texas")) |>
  transmute(
    County = NAME,
    `Poverty Rate` = sprintf("%.2f%%", pov_rate * 100),
    MOE = sprintf("±%.2f pp", pov_moe * 100),
    `Lower Bound` = sprintf("%.2f%%", (pov_rate - pov_moe) * 100),
    `Upper Bound` = sprintf("%.2f%%", (pov_rate + pov_moe) * 100)
  )

knitr::kable(
  uncertainty_pair,
  caption = "Uncertainty in Poverty Rate Estimates"
)
```

Dimmit County has an estimated poverty rate of 41.11% with a margin of error of ±8.88 percentage points, while Zapata County has an estimated rate of 38.39% with a margin of error of ±6.04 percentage points. Their confidence intervals overlap substantially. Therefore, although Dimmit County ranks above Zapata County based on the point estimates, the difference should not be interpreted as a clearly established difference in poverty rates because of the uncertainty in the ACS estimates.

## Methods Note

I used the 2024 ACS 5-year estimates for Texas counties. I used median household income (B19013_001) for the map and calculated the poverty rate using the number of people below poverty (B17001_002) divided by the poverty universe (B17001_001). I also kept the margins of error because ACS estimates include sampling uncertainty. The poverty-rate margin of error was calculated using moe_prop(). The results show that county rankings should be interpreted carefully because counties with similar estimates may have overlapping confidence intervals.