---
title: "Assignment 3: Anatomy of a Chart"
format:
  html:
    code-fold: true
execute:
  warning: false
  message: false
---

## 1. Document a Chart

I selected the boxplot example from Murrell's R graphics code. The comments below describe what each call contributes to the finished figure.

```{r}
# Set the margins for the boxplot
par(mar = c(3, 4.1, 2, 0))

# Draw boxplots for the VC supplement groups at each dose
boxplot(
  len ~ dose,
  data = ToothGrowth,
  boxwex = 0.25,
  at = 1:3 - 0.2,
  subset = supp == "VC",
  col = "white",
  xlab = "",
  ylab = "tooth length",
  ylim = c(0, 35)
)

# Add the x-axis label below the plot
mtext(
  "Vitamin C dose (mg)",
  side = 1,
  line = 2.5,
  cex = 0.8
)

# Add boxplots for the OJ supplement groups beside the VC boxplots
boxplot(
  len ~ dose,
  data = ToothGrowth,
  add = TRUE,
  boxwex = 0.25,
  at = 1:3 + 0.2,
  subset = supp == "OJ"
)

# Add a legend identifying the two supplement groups
legend(
  1.5,
  9,
  c("Ascorbic acid", "Orange juice"),
  fill = c("white", "gray"),
  bty = "n"
)

# Restore the default plot margins
par(mar = c(5.1, 4.1, 4.1, 2.1))
```

## 2. Anscombe's Quartet

The four Anscombe datasets have nearly identical regression statistics, but their plots show very different patterns.

```{r}
data(anscombe)

par(mfrow = c(2, 2))

plot_colors <- c(
  "steelblue",
  "firebrick",
  "darkgreen",
  "purple"
)

plot_symbols <- c(
  16,
  17,
  15,
  18
)

line_types <- c(
  1,
  2,
  3,
  4
)

for(i in 1:4) {

  x <- anscombe[[paste0("x", i)]]
  y <- anscombe[[paste0("y", i)]]

  plot(
    x,
    y,
    main = paste("Dataset", i),
    xlab = paste0("x", i),
    ylab = paste0("y", i),
    pch = plot_symbols[i],
    col = plot_colors[i]
  )

  abline(
    lm(y ~ x),
    col = plot_colors[i],
    lty = line_types[i],
    lwd = 2
  )
}

par(mfrow = c(1, 1))
```

### Regression Comparison

```{r}
results <- data.frame(
  Dataset = 1:4,
  Intercept = NA,
  Slope = NA,
  R_squared = NA,
  Residual_SE = NA
)

for(i in 1:4) {

  x <- anscombe[[paste0("x", i)]]
  y <- anscombe[[paste0("y", i)]]

  model <- lm(y ~ x)

  results$Intercept[i] <- coef(model)[1]
  results$Slope[i] <- coef(model)[2]
  results$R_squared[i] <- summary(model)$r.squared
  results$Residual_SE[i] <- summary(model)$sigma
}

results
```

Although the four datasets have nearly identical slopes, intercepts, R-squared values, and residual standard errors, the visual patterns are very different. Summary statistics alone can hide nonlinear relationships, outliers, and unusual data structures that become clear when the data are plotted.

## 3. Finetune Without Leaving Base R

I used a serif font, non-default colours, and a non-default plotting character.

```{r}
x <- anscombe$x1
y <- anscombe$y1

plot(
  x,
  y,
  main = "Anscombe Dataset 1",
  xlab = "x1",
  ylab = "y1",
  family = "serif",
  col = "steelblue",
  pch = 17,
  cex = 1.2
)

abline(
  lm(y ~ x),
  col = "firebrick",
  lwd = 2
)
```

## 4. Reproduce the Tuned Chart with ggplot2

```{r}
library(ggplot2)

anscombe1 <- data.frame(
  x = anscombe$x1,
  y = anscombe$y1
)

ggplot(
  anscombe1,
  aes(x = x, y = y)
) +
  geom_point(
    color = "steelblue",
    shape = 17,
    size = 3
  ) +
  geom_smooth(
    method = "lm",
    se = FALSE,
    color = "firebrick",
    linewidth = 1
  ) +
  labs(
    title = "Anscombe Dataset 1",
    x = "x1",
    y = "y1"
  ) +
  theme_classic(
    base_family = "serif"
  )
```

The serif font became a `theme()` setting through `theme_classic(base_family = "serif")`. The fixed point colour, shape, and size were specified directly in `geom_point()` rather than through `scale_*()` functions because they were not mapped to data variables. Therefore, none of the settings in this particular chart required a `scale_*()` function. The regression line added with `abline()` in base R was reproduced with `geom_smooth(method = "lm")`. All of the tuned base-R settings used here had a practical ggplot2 equivalent.

## 5. Pre-Hackathon: Daily COVID Deaths in Europe

The target chart shows one point per country per day, a rotated date axis, a single non-default colour, and labels for selected high points.

```{r}
url <- paste0(
  "https://raw.githubusercontent.com/owid/covid-19-data/",
  "master/public/data/owid-covid-data.csv"
)

if(!file.exists("owid-covid-data.csv")) {
  download.file(
    url,
    "owid-covid-data.csv"
  )
}

owidall <- read.csv(
  "owid-covid-data.csv"
)

owidall <- owidall[
  !grepl("^OWID", owidall$iso_code),
]

owideu <- subset(
  owidall,
  continent == "Europe"
)

owideu$date <- as.Date(
  owideu$date
)

covid_plot <- subset(
  owideu,
  !is.na(new_deaths) &
    !is.na(date) &
    new_deaths >= 0
)
```

### Select Points to Label

```{r}
pick_high_point <- function(country, start_date, end_date) {

  d <- subset(
    covid_plot,
    location == country &
      date >= as.Date(start_date) &
      date <= as.Date(end_date)
  )

  d[which.max(d$new_deaths), ]
}

label_points <- rbind(

  pick_high_point(
    "Spain",
    "2020-03-01",
    "2020-04-15"
  ),

  pick_high_point(
    "Spain",
    "2020-04-16",
    "2020-06-30"
  ),

  pick_high_point(
    "Germany",
    "2020-10-01",
    "2021-03-31"
  ),

  pick_high_point(
    "Germany",
    "2022-07-01",
    "2023-04-30"
  ),

  pick_high_point(
    "Ukraine",
    "2021-07-01",
    "2022-03-31"
  ),

  pick_high_point(
    "Italy",
    "2023-01-01",
    "2023-08-31"
  )
)
```

### Replicated Chart

```{r}
par(mfrow = c(1, 1))
par(mar = c(7, 5, 3, 2))

plot_start <- as.Date("2020-01-01")
plot_end <- as.Date("2023-08-31")

plot(
  covid_plot$date,
  covid_plot$new_deaths,
  pch = 16,
  cex = 0.45,
  col = "deeppink",
  xlab = "",
  ylab = "COVID Deaths in Europe (Daily)",
  xaxt = "n",
  xlim = c(plot_start, plot_end),
  ylim = c(
    0,
    max(
      covid_plot$new_deaths[
        covid_plot$date >= plot_start &
          covid_plot$date <= plot_end
      ],
      na.rm = TRUE
    ) * 1.12
  )
)

axis.Date(
  1,
  at = seq(
    plot_start,
    plot_end,
    length.out = 12
  ),
  format = "%Y-%m-%d",
  las = 2,
  cex.axis = 0.7
)

mtext(
  "Date",
  side = 1,
  line = 5.5
)

text(
  label_points$date,
  label_points$new_deaths,
  labels = label_points$location,
  pos = 3,
  offset = 0.4,
  cex = 0.65
)

par(
  mar = c(5.1, 4.1, 4.1, 2.1)
)
```

## AI Use Appendix

ChatGPT was used for troubleshooting selected R errors and clarifying parts of the assignment instructions. I executed the code, reviewed the outputs, and made the final decisions about the figures and content.