---
title: "Assignment 2"
format:
  html:
    code-fold: true
execute:
  warning: false
  message: false
---

## Murrell's R Graphics Chapter 1

I worked through the Chapter 1 code from Murrell's *R Graphics* line by line and observed the changes produced by each graphics call.

[View the Chapter 1 practice code](murrell01_practice.R)

## Happy Planet Index Data

I selected 2025 because it is the most recent year available in the dataset.

```{r}
library(readxl)

hpi <- read_excel(
  "Happy-Planet-Index-2006-2025-public-data-set.xlsx",
  sheet = "All Data",
  skip = 1
)

hpi_2025 <- subset(hpi, Year == 2025)

par(mfrow = c(1, 1), pty = "m")
```

## 1. `par()`

`par()` controls graphical parameters, such as the margins and layout of a plot.

```{r}
par(mar = c(5, 5, 3, 2))

plot(
  hpi_2025$`Life Expectancy`,
  hpi_2025$`Life Satisfaction`,
  xlab = "Life Expectancy",
  ylab = "Life Satisfaction",
  main = "Life Expectancy and Life Satisfaction, 2025",
  pch = 16
)
```

## 2. `box()`

`box()` adds a box around the current plot.

```{r}
plot(
  hpi_2025$`Life Expectancy`,
  hpi_2025$`Life Satisfaction`,
  xlab = "Life Expectancy",
  ylab = "Life Satisfaction",
  main = "Life Expectancy and Life Satisfaction, 2025",
  pch = 16,
  axes = FALSE
)

axis(1)
axis(2)
box()
```

## 3. `axis()`

`axis()` adds or customizes an axis on the current plot.

```{r}
plot(
  hpi_2025$`Life Expectancy`,
  hpi_2025$`Life Satisfaction`,
  xlab = "",
  ylab = "",
  main = "Life Expectancy and Life Satisfaction, 2025",
  pch = 16,
  axes = FALSE
)

axis(
  side = 1,
  at = seq(55, 85, by = 5)
)

axis(
  side = 2,
  at = seq(3, 8, by = 1)
)

box()

mtext("Life Expectancy", side = 1, line = 3)
mtext("Life Satisfaction", side = 2, line = 3)
```

## 4. `hist()`

`hist()` displays the distribution of a numeric variable by grouping values into intervals.

```{r}
hist(
  hpi_2025$`Life Expectancy`,
  main = "Distribution of Life Expectancy, 2025",
  xlab = "Life Expectancy",
  ylab = "Frequency"
)
```

## 5. `boxplot()`

`boxplot()` summarizes the distribution of a numeric variable using the median, quartiles, and potential outliers.

```{r}
boxplot(
  hpi_2025$`Life Satisfaction`,
  main = "Distribution of Life Satisfaction, 2025",
  ylab = "Life Satisfaction"
)
```

## 6. `pie()`

`pie()` displays the proportions of categories as slices of a circle.

```{r}
continent_names <- c(
  "Latin America",
  "N America & Oceania",
  "Western Europe",
  "Middle East & N. Africa",
  "Africa",
  "South Asia",
  "Eastern Europe & Central Asia",
  "East Asia"
)

continent_counts <- table(hpi_2025$Continent)

pie(
  continent_counts,
  labels = continent_names,
  main = "Countries by Continent, 2025"
)
```

## 7. `persp()`

`persp()` creates a three-dimensional perspective plot of a surface.

```{r}
le_group <- cut(
  hpi_2025$`Life Expectancy`,
  breaks = 5,
  labels = FALSE
)

ls_group <- cut(
  hpi_2025$`Life Satisfaction`,
  breaks = 5,
  labels = FALSE
)

z_matrix <- tapply(
  hpi_2025$`Ecological Footprint`,
  list(le_group, ls_group),
  mean,
  na.rm = TRUE
)

z_matrix[is.na(z_matrix)] <- mean(
  hpi_2025$`Ecological Footprint`,
  na.rm = TRUE
)

par(mar = c(2, 2, 3, 2))

persp(
  x = 1:5,
  y = 1:5,
  z = z_matrix,
  theta = 30,
  phi = 30,
  expand = 0.8,
  ticktype = "detailed",
  xlab = "Life Expectancy Group",
  ylab = "Life Satisfaction Group",
  zlab = "Mean Ecological Footprint",
  main = "Mean Ecological Footprint, 2025"
)
```

## 8. `points()`

`points()` adds points to an existing plot.

```{r}
par(mar = c(5, 4, 4, 2) + 0.1)

plot(
  hpi_2025$`Life Expectancy`,
  hpi_2025$`Life Satisfaction`,
  xlab = "Life Expectancy",
  ylab = "Life Satisfaction",
  main = "Life Expectancy and Life Satisfaction, 2025",
  pch = 1
)

high_life <- hpi_2025$`Life Expectancy` >= 80

points(
  hpi_2025$`Life Expectancy`[high_life],
  hpi_2025$`Life Satisfaction`[high_life],
  pch = 16
)
```

## 9. `lines()`

`lines()` adds lines to an existing plot.

```{r}
x <- hpi_2025$`Life Expectancy`
y <- hpi_2025$`Life Satisfaction`

plot(
  x,
  y,
  xlab = "Life Expectancy",
  ylab = "Life Satisfaction",
  main = "Life Expectancy and Life Satisfaction, 2025",
  pch = 16
)

fit <- lm(y ~ x)

x_line <- seq(
  min(x, na.rm = TRUE),
  max(x, na.rm = TRUE),
  length.out = 100
)

y_line <- predict(
  fit,
  newdata = data.frame(x = x_line)
)

lines(
  x_line,
  y_line,
  lwd = 2
)
```

## 10. `text()`

`text()` adds text labels at specified coordinates on an existing plot.

```{r}
plot(
  hpi_2025$`Life Expectancy`,
  hpi_2025$`Life Satisfaction`,
  xlab = "Life Expectancy",
  ylab = "Life Satisfaction",
  main = "Life Expectancy and Life Satisfaction, 2025",
  pch = 16,
  cex = 0.65,
  col = "black",
  ylim = c(3, 8.3)
)

top_country <- which.max(hpi_2025$`Life Satisfaction`)

points(
  hpi_2025$`Life Expectancy`[top_country],
  hpi_2025$`Life Satisfaction`[top_country],
  pch = 16,
  cex = 1.2,
  col = "firebrick3"
)

text(
  hpi_2025$`Life Expectancy`[top_country],
  hpi_2025$`Life Satisfaction`[top_country],
  labels = hpi_2025$Country[top_country],
  pos = 2,
  offset = 0.8,
  col = "firebrick3",
  font = 2,
  cex = 0.9
)
```

## 11. `mtext()`

`mtext()` adds text to the margins of a plot.

```{r}
par(mar = c(5, 4, 4, 2) + 0.1)

plot(
  hpi_2025$`Life Expectancy`,
  hpi_2025$`Life Satisfaction`,
  xlab = "Life Expectancy",
  ylab = "Life Satisfaction",
  main = "Life Expectancy and Life Satisfaction, 2025",
  pch = 16,
  cex = 0.65
)

mtext(
  "Happy Planet Index data, 2025",
  side = 1,
  line = 4,
  cex = 0.8
)
```

## 12. `legend()`

`legend()` adds a key that explains symbols or groups used in a plot.

```{r}
plot(
  hpi_2025$`Life Expectancy`,
  hpi_2025$`Life Satisfaction`,
  xlab = "Life Expectancy",
  ylab = "Life Satisfaction",
  main = "Life Expectancy and Life Satisfaction, 2025",
  pch = 1
)

high_life <- hpi_2025$`Life Expectancy` >= 80

points(
  hpi_2025$`Life Expectancy`[high_life],
  hpi_2025$`Life Satisfaction`[high_life],
  pch = 16
)

legend(
  "topleft",
  legend = c(
    "Life Expectancy < 80",
    "Life Expectancy >= 80"
  ),
  pch = c(1, 16),
  bty = "n"
)
```
