---
title: "Assignment 4: 48-Hour Chart Hackathon"
format: html
---

## Setup

```{r}
#| warning: false
#| message: false

library(readxl)
library(tidyverse)
library(patchwork)

hpi <- read_excel(
  "Happy-Planet-Index-2006-2025-public-data-set.xlsx",
  sheet = "All Data",
  skip = 1
)

hpi_2025 <- hpi %>%
  filter(Year == 2025)
```

```{r}
#| include: false

names(hpi)
nrow(hpi_2025)
unique(hpi_2025$Year)
summary(hpi_2025)
```

## Chart 1: Variable-Width Column

The 2025 HPI data were summarized by continent. The width of each column represents total population, while the height represents mean HPI across countries within each continent.

```{r}
hpi_continent <- hpi_2025 %>%
  group_by(Continent) %>%
  summarise(
    population = sum(Population, na.rm = TRUE),
    mean_hpi = mean(HPI, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    continent_name = case_when(
      Continent == 1 ~ "Latin America",
      Continent == 2 ~ "N America & Oceania",
      Continent == 3 ~ "Western Europe",
      Continent == 4 ~ "Middle East & N. Africa",
      Continent == 5 ~ "Africa",
      Continent == 6 ~ "South Asia",
      Continent == 7 ~ "Eastern Europe & Central Asia",
      Continent == 8 ~ "East Asia"
    )
  ) %>%
  arrange(desc(mean_hpi)) %>%
  mutate(
    xmax = cumsum(population),
    xmin = lag(xmax, default = 0),
    xmid = (xmin + xmax) / 2,
    continent_name = factor(
      continent_name,
      levels = continent_name
    )
  )
```

```{r}
#| include: false

hpi_continent %>%
  select(continent_name, population, mean_hpi, xmin, xmax)
```

```{r}
#| fig-width: 7
#| fig-height: 4.8

ggplot(hpi_continent) +
  geom_rect(
    aes(
      xmin = xmin,
      xmax = xmax,
      ymin = 0,
      ymax = mean_hpi,
      fill = continent_name
    ),
    color = "white",
    linewidth = 0.4
  ) +

  geom_text(
    aes(
      x = xmid,
      y = mean_hpi + 1.2,
      label = sprintf("%.1f", mean_hpi)
    ),
    size = 2.6
  ) +

  scale_fill_manual(
    values = c(
      "#0072B2",
      "#E69F00",
      "#009E73",
      "#CC79A7",
      "#D55E00",
      "#56B4E9",
      "#F0E442",
      "#999999"
    )
  ) +

  scale_x_continuous(
    labels = scales::label_number(
      scale = 1e-6,
      suffix = "B",
      accuracy = 0.1
    ),
    expand = c(0, 0)
  ) +

  scale_y_continuous(
    limits = c(0, 64),
    breaks = seq(0, 60, 10),
    expand = c(0, 0)
  ) +

  labs(
    title = "Latin America Has the Highest Mean HPI in 2025",
    subtitle = "Column width represents population; height represents mean HPI",
    x = "Population (billions)",
    y = "Mean HPI",
    fill = NULL,
    caption = "Source: Happy Planet Index"
  ) +

  guides(
    fill = guide_legend(
      ncol = 1,
      keywidth = grid::unit(0.38, "cm"),
      keyheight = grid::unit(0.28, "cm")
    )
  ) +

  theme_minimal() +

  theme(
    plot.title = element_text(
      size = 13,
      face = "bold",
      hjust = 0,
      margin = margin(b = 3)
    ),

    plot.subtitle = element_text(
      size = 8,
      margin = margin(b = 7)
    ),

    axis.title = element_text(size = 9.5),
    axis.text = element_text(size = 8.5),

    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),

    legend.position = "right",
    legend.justification = "center",
    legend.text = element_text(size = 7.8),
    legend.spacing.y = grid::unit(0.07, "cm"),
    legend.margin = margin(l = 7),

    plot.caption = element_text(
      size = 7.5,
      hjust = 1,
      margin = margin(t = 4)
    ),

    plot.margin = margin(
      t = 5,
      r = 5,
      b = 4,
      l = 5
    )
  )
```

## Chart 3: Bar Chart

The 2025 HPI data were used to identify the 20 countries with the highest scores and the 20 countries with the lowest scores. Countries are sorted by HPI within each group.

```{r}
hpi_ranked <- hpi_2025 %>%
  filter(!is.na(HPI))

top20 <- hpi_ranked %>%
  slice_max(
    order_by = HPI,
    n = 20,
    with_ties = FALSE
  )

bottom20 <- hpi_ranked %>%
  slice_min(
    order_by = HPI,
    n = 20,
    with_ties = FALSE
  )
```

```{r}
#| include: false

nrow(top20)
nrow(bottom20)

top20 %>%
  select(Country, HPI) %>%
  arrange(desc(HPI))

bottom20 %>%
  select(Country, HPI) %>%
  arrange(desc(HPI))
```

```{r}
#| fig-width: 12
#| fig-height: 7
#| fig-align: center

p_top <- ggplot(
  top20,
  aes(
    x = HPI,
    y = forcats::fct_reorder(Country, HPI)
  )
) +
  geom_col(
    fill = "#0072B2",
    width = 0.72
  ) +
  geom_text(
    aes(
      label = sprintf("%.1f", HPI)
    ),
    hjust = -0.15,
    size = 2.5
  ) +
  scale_x_continuous(
    limits = c(0, 72),
    breaks = seq(0, 70, 10),
    expand = c(0, 0)
  ) +
  labs(
    title = "Top 20",
    x = "Happy Planet Index",
    y = NULL
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(
      size = 11,
      face = "bold",
      margin = margin(b = 5)
    ),
    axis.title.x = element_text(size = 9.5),
    axis.text = element_text(size = 8),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  )

p_bottom <- ggplot(
  bottom20,
  aes(
    x = HPI,
    y = forcats::fct_reorder(Country, HPI)
  )
) +
  geom_col(
    fill = "#D55E00",
    width = 0.72
  ) +
  geom_text(
    aes(
      label = sprintf("%.1f", HPI)
    ),
    hjust = -0.15,
    size = 2.5
  ) +
  scale_x_continuous(
    limits = c(0, 72),
    breaks = seq(0, 70, 10),
    expand = c(0, 0)
  ) +
  labs(
    title = "Bottom 20",
    x = "Happy Planet Index",
    y = NULL
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(
      size = 11,
      face = "bold",
      margin = margin(b = 5)
    ),
    axis.title.x = element_text(size = 9.5),
    axis.text = element_text(size = 8),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  )

p_top + p_bottom +
  plot_annotation(
    title = "A Clear HPI Gap Separates the Top and Bottom 20 Countries in 2025",
    subtitle = "The lowest score in the Top 20 is 57.6, compared with 41.8 at the top of the Bottom 20",
    caption = "Source: Happy Planet Index",
    theme = theme(
      plot.title = element_text(
        size = 14,
        face = "bold",
        margin = margin(b = 3)
      ),
      plot.subtitle = element_text(
        size = 9,
        margin = margin(b = 7)
      ),
      plot.caption = element_text(
        size = 7.5,
        hjust = 1,
        margin = margin(t = 5)
      )
    )
  )
```